---
title: "document1"
output:
  pdf_document: default
  html_document: 
    theme: cosmo
date: "2025-01-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Just type as always.Example for Markdown.

Second paragraph.

Leslie Paredes  
Department of Biology  
University of Vermont  

# First header (largest)

Back to plaint text here 
 
## Second level

### Thrid level

use asterisks for *italic* and double asterisks for **boldface** text. Use caret for ^superscript^ and ~subscript~ and two tildas for ~~strikethroughh~~

>use to geater than sign for indented quotesd material
Looks nice!

## For lists

- first item
- second item
  + indented list item

links [linked phrase](website address)

First Header | Second Header
------------ | -------------
Content cell | Content cell
Content cell | Content cell

### Fencing options

Using a single tick for ´in line fencing´ of material.

Use triple back ticks for a fenced block of text.

```
everything here is plain text
even with markdown

properly spaced
```
Use r in an in-line text:

the value of pi is `r pi +10`

## Use chunks for r code

```{r, eval=TRUE,echo=TRUE}
# r code can be used here!
print(runif(10))
plot(runif(10),runif(10))

```
### Using LaTex for math

Use a single dollar sign $a = b + c$ to insert mathematical statements within sentences of plain text.

In contrast, use a double dollar sign: 

$$a = b +c$$
to center and offset equations.

$$H_0 = Z_{a + b}$$
$$S = cA^z$$
$$S=cA^z_1 + z_{2 + x}$$
$$\alpha = \frac{\beta}{\delta + \gamma_x}$$
$$z = \sum_{i=1}^X{K}$$
$$\backslash$$
$$\backslash \alpha \le b \backslash$$
$$P(Occurrence of Species A) = Z$$
$$P(\mbox{Occurrence of Species A}) = Z$$
```{r, echo=TRUE, eval=TRUE}
#here is a new chunk of code, distant from the first one 

z<-1:10
print(z)
#end of the second chunk
```

